---
title: "Final_Project: Cereals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
#install.packages("treemap")
library(treemap)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2) 
cereal <- read_csv("DATA 2401/cereal.csv", na = "-1")
glimpse(cereal)
```
# Types of cereal: Hot vs Cold
```{r}
cereal %>% 
  count(type)

```
- What are the 3 hot cereals
```{r}
cereal %>% 
  select(name, type) %>% 
  filter(type == "H")
```
# ounces to grams:
```{r}
cereal <- cereal %>% 
  mutate(weight_grams = round(weight*28.3495, digits = 0), weight_ounces = weight, rating = ceiling(rating)) %>% 
  subset(select = -weight)
glimpse(cereal)
```

# Which cereal has the highest calories per serving?
 - highest is the Muesilx Crispy Blend with 160 calories per serving

```{r,  }
cereal$name[cereal$calories == max(cereal$calories)]

cereal %>% 
  select(name, calories, fat, carbo, sugars, protein) %>% 
  arrange(desc(calories)) %>%
  head(5) %>% 
  ggplot(aes(x = calories, y = name)) + 
  geom_col() +
  labs(title = "Top 5 Cereals with Highest Calories per Serving", x = "Calories", y= "Name")


```

# Which cereals have the lowest calories per serving?
  - A three way tie between Puffed Wheat, Puffed Rice, and All- Bran with extra fiber
```{r}
cereal$name[cereal$calories == min(cereal$calories)]

cereal %>% 
  select(name, type, calories) %>% 
  arrange(calories) %>% 
  head(5) %>% 
  ggplot(aes(x= calories, y = name)) +
  geom_col() +
  labs(title = "5 Cereals with the worst Calories per Serving", x = "Calories", y= "Name")

```

# Which has the highest protein per serving?
```{r}
cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5)

cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5) %>% 
  ggplot(aes(x = name, y = protein))+
  geom_col() +
  labs(title = "5 Cereals with the most amount of protein")

```


# ounces to grams:
```{r}
cereal <- cereal %>% 
  mutate(weight_grams = round(weight*28.3495, digits = 0), weight_ounces = weight, rating = ceiling(rating)) %>% 
  subset(select = -weight)
cereal
```

# Which cereals have the lowest calories per serving?
```{r}
cereal %>% 
  select(name, type, calories) %>% 
  arrange(calories) %>% 
  head(10) %>% 
  ggplot(aes(x= calories, y = name)) +
  geom_col()
```

# Which has the highest protein per serving?
```{r}
cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5)

cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5) %>% 
  ggplot(aes(x = name, y = protein))+
  geom_col()

```

# Do cereals with the highest sugar content have more calories? 
```{r}
cereal %>% 
  select(name, sugars, calories) %>% 
  arrange(desc(sugars))
cereal %>% 
  arrange(desc(sugars)) %>% 
  ggplot(aes(x = calories, y = sugars)) +
  geom_point() +
  geom_smooth()  

  
```

# What factors determine the ratings of the cereals? What affects the ratings? Does the shelf display affect the ratings?  Which self display(1, 2, or 3) have the highest/lowest ratings? 

```{r}
#treemap(cereal, index = c("rating"), vSize = "sugars")
topR <-cereal %>% 
  arrange(desc(rating)) %>% 
  head(5) 
ave_top_cal <- mean(topR$calories)
ave_top_pro <- mean(topR$protein)
ave_top_fib <- mean(topR$fiber)
ave_top_sug <- mean(topR$sugars)
ave_top_sod <- mean(topR$sodium)
ave_top_pot <- mean(topR$potass)
ave_top_vit <- mean(topR$vitamins)


bottomostR <- cereal %>% 
  arrange(rating) %>% 
  head(5) 
ave_bottom_cal <- mean(bottomostR$calories)
ave_bottom_pro <- mean(bottomostR$protein)
ave_bottom_fib <- mean(bottomostR$fiber)
ave_bottom_sug <- mean(bottomostR$sugars)
ave_bottom_sod <- mean(bottomostR$sodium)
ave_bottom_pot <- mean(bottomostR$potass)
ave_bottom_vit <- mean(bottomostR$vitamins)



topR
bottomostR
```

#Which has the most fiber? Will high fiber cereals have lower calories per serving?  
```{r}

cereal %>% 
  select(name, fiber, calories) %>% 
  arrange(desc(fiber)) %>% 
  head(5)

cereal %>% 
  arrange(desc(fiber)) %>% 
  head(5) %>% 
  ggplot(aes(x = calories, y = fiber, fill = name)) +
  geom_col() +
  theme_classic()


```

# Are there any keto-friendly cereals? Which are low-carb? 
```{r}
cereal %>% 
  filter(type == "H") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)
cereal %>% 
  filter(type == "C") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)

cereal %>% 
  filter(type == "C") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo) %>% 
  head(5) %>% 
  ggplot(aes(x= calories, y= carbo, fill = name))+
  geom_label(aes(label = name))
```
# What are the Calories grouped by shelfs?
```{r}
cereal %>% 
  ggplot(aes(calories, shelf, group = shelf, fill = shelf)) + 
  geom_violin()

```
# Which self display(1, 2, or 3) have the highest/lowest ratings?
cereal %>% 
  filter(shelf == 1) %>% 
  ggplot(aes(x = reorder(name, rating), y = rating, fill = rating ))+
  geom_col() + 
  coord_flip()
cereal %>% 
  filter(shelf == 2) %>% 
  ggplot(aes(x = reorder(name, rating), y = rating, fill = rating ))+
  geom_col() + 
  coord_flip()
cereal %>% 
  filter(shelf == 3) %>% 
  ggplot(aes(x = reorder(name, rating), y = rating, fill = rating ))+
  geom_col() + 
  coord_flip()
