---
title: "Final_Project: Cereals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2) 
cereal <- read_csv("DATA 2401/cereal.csv", na = "-1")
glimpse(cereal)
```
# Types of cereal: Hot vs Cold
```{r}
cereal %>% 
  count(type)
```


# ounces to grams:
```{r}
cereal <- cereal %>% 
  mutate(weight_ounces = weight, weight_grams = round(weight*28.3495, 0)) %>% 
  subset(select = -weight)
cereal
```


#Which cereal has the highest calories per serving?

```{r }
cereal %>% 
  filter(calories >= 100, cups >= 1) %>% 
  select(name, calories, cups) 
```

#Which cereal has the lowest calories per serving?
```{r}
cereal %>% 
  select(name, calories) %>% 
  summarize(min(calories)) 
cereal %>% 
  select(name, type, calories) %>% 
  arrange(calories) 
```

#Which has the highest protein per serving?
```{r}

cereal %>% 
  select(name, type, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5)

```

#Do higher-sugar cereals have more calories? 
```{r}
cereal %>% 
  select(name, calories, sugars) %>% 
  arrange(desc(sugars))

  
```

#Does the shelf display affect the ratings?  Which self display(1, 2, or 3) have the highest/lowest ratings? 
```{r}
colnames(cereal)
cereal %>% 
  select(name, rating, shelf) %>% 
  arrange(desc(rating))
```

#Which has the most fiber? Will high fiber cereals have lower calories per serving?  
```{r}
cereal %>% 
  select(name, fiber, calories) %>% 
  arrange(desc(fiber)) %>% 
  head(10)

```

#Which are low-carb? Are there any keto-friendly cereals? 
```{r}
cereal %>% 
  filter(type == "H") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)
cereal %>% 
  filter(type == "C") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)
  
```


