---
title: "Final_Project: Cereals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2) 
cereal <- read_csv("DATA 2401/cereal.csv", na = "-1")
glimpse(cereal)
```
# Types of cereal: Hot vs Cold
```{r}
cereal %>% 
  count(type)

```


# ounces to grams:
```{r}
cereal <- cereal %>% 
  mutate(weight_grams = round(weight*28.3495, digits = 0), weight_ounces = weight, rating = ceiling(rating)) %>% 
  subset(select = -weight)
cereal
```

#Which cereal has the highest calories per serving?

```{r,  }
cereal %>% 
  select(name, calories, fat, carbo, sugars, protein) %>% 
  arrange(desc(calories)) %>%
  head(5) %>% 
  ggplot(aes(x = calories, y = name)) + 
  geom_col() +
  labs(title = "Top 5 Cereals with Highest Calories per Serving", x = "Calories", y= "Name")


```

#Which cereals have the lowest calories per serving?
```{r}
cereal %>% 
  select(name, type, calories) %>% 
  arrange(calories) %>% 
  head(10) %>% 
  ggplot(aes(x= calories, y = name)) +
  geom_col()
```

#Which has the highest protein per serving?
```{r}
cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5)

cereal %>% 
  select(name, calories, protein) %>% 
  arrange(desc(protein)) %>% 
  head(5) %>% 
  ggplot(aes(x = name, y = protein))+
  geom_col()

```

#Do cereals with the highest sugar content have more calories? 
```{r}
#cereal %>% 
 # select(name, sugars, calories) %>% 
# arrange(desc(sugars))
cereal %>% 
  arrange(desc(sugars)) %>% 
  ggplot(aes(x = calories, y = sugars)) +
  geom_smooth() +
  geom_point()

  
```

# What factors determine the ratings of the cereals? What affects the ratings? Does the shelf display affect the ratings?  Which self display(1, 2, or 3) have the highest/lowest ratings? 
```{r}

cereal %>% 
  arrange(desc(rating)) %>% 
  head(5)

cereal %>% 
  arrange(rating) %>% 
  head(5)


```

#Which has the most fiber? Will high fiber cereals have lower calories per serving?  
```{r}

cereal %>% 
  select(name, fiber, calories) %>% 
  arrange(desc(fiber)) %>% 
  head(5)

cereal %>% 
  arrange(desc(fiber)) %>% 
  head(5) %>% 
  ggplot(aes(x = calories, y = fiber, fill = name)) +
  geom_col() +
  theme_classic()


```

# Are there any keto-friendly cereals? Which are low-carb? 
```{r}
cereal %>% 
  filter(type == "H") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)
cereal %>% 
  filter(type == "C") %>% 
  select(name, type, calories, carbo) %>% 
  arrange(carbo)
```


